---
title: "STAT/MATH 495: Problem Set 02"
author: "Anthony Rentsch"
date: "2017-09-19"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 2
    collapsed: false
    smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.width=8, fig.height=4.5)

# Load packages
library(tidyverse)
library(broom)
library(corrplot)

# Note the relative file path, and not absolute file path:
# http://www.coffeecup.com/help/articles/absolute-vs-relative-pathslinks/
train <- read.csv("data/train.csv")
test <- read.csv("data/test.csv")
macro <- read.csv("data/macro.csv")
```


# Exploratory Data Analysis

Visualizing the distribution of sale price using a histogram, we can see the data has a strong right skew. We might consider a transformation to make sale price more normally distributed. However, for the sake of model interpretability, we chose to proceed with sale price untransformed.
```{r}
ggplot(data = train) + geom_histogram(aes(x = price_doc), fill = "cadetblue3", bins = 60, color = "cadetblue4")+ ggtitle("Histogram of Home Prices") + 
  xlab("Sale Price") +
  ylab("Count") +
  theme(plot.title = element_text(hjust = 0.5))
```

Visualize correlation between all structure describing variables with high availability in the data set, excluding observations with NA values.
```{r}
new <- train[,c(3,4,5,6,8,9,10,292)]
M <- cor(new, use="pairwise.complete.obs")
corrplot(M, method = "circle")
```

The num_rooms variable looks like it has strongest linear correlation with price_doc so we chose num_rooms to create our splines model.
```{r}
ggplot(data = train) + geom_point(aes(x = num_room, y = price_doc)) + 
  labs(title = "Scatterplot of number of rooms vs. price") +
  theme(plot.title = element_text(hjust = 0.5))
```

# Model Fit

Remove properties for which the number of rooms is not included.
```{r}
new.train <- train[-which(is.na(train$num_room)),]
```

Fit spline models for four different degress of freedom.
```{r}
two <- smooth.spline(new.train$num_room, new.train$price_doc, df=2) %>%
  broom::augment() %>% 
  ggplot(aes(x = new.train$num_room)) +
  geom_point(aes(y = new.train$price_doc)) +
  geom_line(aes(y=.fitted), col="blue", size=1) +
  labs(title = "2 degrees of freedom", x = "num_room", y = "price_doc") +
  theme(plot.title = element_text(hjust = 0.5))

seven <- smooth.spline(new.train$num_room, new.train$price_doc, df=7) %>%
  broom::augment() %>% 
  ggplot(aes(x = new.train$num_room)) +
  geom_point(aes(y = new.train$price_doc)) +
  geom_line(aes(y=.fitted), col="blue", size=1) +
  labs(title = "7 degrees of freedom", x = "num_room", y = "price_doc") +
  theme(plot.title = element_text(hjust = 0.5))

 ten <- smooth.spline(new.train$num_room, new.train$price_doc, df=10) %>%
  broom::augment() %>% 
  ggplot(aes(x = new.train$num_room)) +
  geom_point(aes(y = new.train$price_doc)) +
  geom_line(aes(y=.fitted), col="blue", size=1) +
  labs(title = "10 degrees of freedom", x = "num_room", y = "price_doc") +
   theme(plot.title = element_text(hjust = 0.5))

thirteen <- smooth.spline(new.train$num_room, new.train$price_doc, df=13) %>%
  broom::augment() %>% 
  ggplot(aes(x = new.train$num_room)) +
  geom_point(aes(y = new.train$price_doc)) +
  geom_line(aes(y=.fitted), col="blue", size=1) +
  labs(title = "13 degrees of freedom", x = "num_room", y = "price_doc") +
  theme(plot.title = element_text(hjust = 0.5))

two
seven
ten
thirteen
```

# Review and Create Submission File

Will use df = 7 to create spline model, as it appears to give best fit based on our plots.
```{r}
spline.model <- smooth.spline(new.train$num_room, new.train$price_doc, df = 7)
```

Predict the price_doc for the test set using spline.model.
```{r}
new_data = test$num_room
pred <- predict(spline.model, new_data)
pred <- as.data.frame(pred)
```

Plot original data, spline model, and predictions.
```{r}
spline.model %>% broom::augment() %>% ggplot() +
  geom_point(aes(x = new.train$num_room, y = new.train$price_doc)) +
  geom_line(aes(x = new.train$num_room, y=.fitted), col="blue", size=1) +
  geom_point(data = pred, aes(x = pred$x, y = pred$y), col = "red", size = 3) +
  labs(x = "num_room", y = "price_doc")
```

Create a clean list with id and predicted price_doc and write this list to a .csv file.
```{r}
pred <- cbind(id = test$id, pred)
pred <- pred[,-2]
colnames(pred) <- c("id","price_doc")

write.csv(pred, file = "submission.csv", row.names = F)
```

# Retry with outliers removed

Our first Kaggle submission yielded a score of .45256. Let’s investigate if we can improve that score by removing the high outliers.

Remove properties with 8 or more rooms. (Value chosen by examing scatter plot of sale price vs number of rooms.)
```{r}
new.train2 <- new.train[-which(new.train$num_room >= 8),]
```

Fit spline models for four different degress of freedom. We’ve used different degrees of freedom here since we cannot use more degrees of freedom than we have x values.
```{r}
two <- smooth.spline(new.train2$num_room, new.train2$price_doc, df=2) %>%
  broom::augment() %>% 
  ggplot(aes(x = new.train2$num_room)) +
  geom_point(aes(y = new.train2$price_doc)) +
  geom_line(aes(y=.fitted), col="blue", size=1) +
  labs(title = "1 degrees of freedom", x = "num_room", y = "price_doc") +
  theme(plot.title = element_text(hjust = 0.5))

three <- smooth.spline(new.train2$num_room, new.train2$price_doc, df=3) %>%
  broom::augment() %>% 
  ggplot(aes(x = new.train2$num_room)) +
  geom_point(aes(y = new.train2$price_doc)) +
  geom_line(aes(y=.fitted), col="blue", size=1) +
  labs(title = "3 degrees of freedom", x = "num_room", y = "price_doc") +
  theme(plot.title = element_text(hjust = 0.5))

 five <- smooth.spline(new.train2$num_room, new.train2$price_doc, df=5) %>%
  broom::augment() %>% 
  ggplot(aes(x = new.train2$num_room)) +
  geom_point(aes(y = new.train2$price_doc)) +
  geom_line(aes(y=.fitted), col="blue", size=1) +
  labs(title = "5 degrees of freedom", x = "num_room", y = "price_doc") +
   theme(plot.title = element_text(hjust = 0.5))

seven <- smooth.spline(new.train2$num_room, new.train2$price_doc, df=7) %>%
  broom::augment() %>% 
  ggplot(aes(x = new.train2$num_room)) +
  geom_point(aes(y = new.train2$price_doc)) +
  geom_line(aes(y=.fitted), col="blue", size=1) +
  labs(title = "7 degrees of freedom", x = "num_room", y = "price_doc") +
  theme(plot.title = element_text(hjust = 0.5))

two
three
five
seven
```

In this case, df = 3 and df = 5 both appear to work well. We settled in the middle on df = 4.
```{r}
spline.model2 <- smooth.spline(new.train2$num_room, new.train2$price_doc, df = 4)
```

Predict the price_doc for the test set using spline.model2.
```{r}
new_data2 = test$num_room
pred2 <- predict(spline.model2, new_data2)
pred2 <- as.data.frame(pred2)
```

Plot original data, spline model, and predictions.
```{r}
spline.model2 %>% broom::augment() %>% ggplot() +
  geom_point(aes(x = new.train2$num_room, y = new.train2$price_doc)) +
  geom_line(aes(x = new.train2$num_room, y=.fitted), col="blue", size=1) +
  geom_point(data = pred2, aes(x = pred2$x, y = pred2$y), col = "red", size = 3) +
  labs(x = "num_room", y = "price_doc")
```

Create a clean list with id and predicted price_doc and write this list to a .csv file for submission to Kaggle.
```{r}
pred2 <- cbind(id = test$id, pred2)
pred2 <- pred2[,-2]
colnames(pred2) <- c("id","price_doc")

write.csv(pred2, file = "submission2.csv", row.names = F)
```

Our new Kaggle score is .45247. So, it appears removing outliers improved our model slightly.